This is an example of article title

### 2025-10-31 paper/paper.tex
Balance the last page.
improve second example

Code~\ref{lst:rod-dynamics} shows functions that calculate the right hand side of the ODE, where \lstinline{E_field} is a function that calculates electric field strength from the point charges at the given point.
The function \lstinline{force_and_torque} calculates the force and torque affecting the rod, while \lstinline{dynamics} converts them to the Lie algebra element can be used by the solver.
\begin{lstlisting}[
    language = Julia,
    numbers=left,
    label={lst:rod-dynamics},
    caption={Dynamics of a rod affected by point charges.},
    mathescape
]
function force_and_torque(
        sp::SystemParameters, Fθ, r
    )
    u, u$\bot$ = eachrow(Fθ)
    F, $\tau$ = zeros(2), 0.0
    for s in range(-sp.L/2, sp.L/2; length=Nq)
        E = E_field(sp, r .+ s .* u)
        F .+= sp.$\lambda$ * E * (sp.L/(Nq-1))
        $\tau$ += sp.$\lambda$ * s * dot(u$\bot$, E) * (sp.L/(Nq-1))
    end
    return F, $\tau$
end
# f((p,θ), (dp,dθ), t) including system parameters
function dynamics(state, sp::SystemParameters, t)
    p, dp = state.x
    F, tau = force_and_torque(sp, p.x...)
    a = F ./ sp.M
    coords = [tau / sp.Icm, a...]
    ddx = hat(se2, coords, ArrayPartition) #TODO se2 from global scope?
    return ArrayPartition(dp, ddx)
end
\end{lstlisting}

@@

\begin{lstlisting}[
    language = Julia,
    numbers=left,
    label={lst:heun},
    caption={Step calculation for the RKMK Heun method.}
]
function integrator_step!(
        A::GroupAction, sp::SystemParameters,
        y, f, t, dt
) # TODO how are f and dynamics are they the same related?
    Ie = Identity(A.group)
    F1 = (dt / 2) * f(y, sp, t)
    tmp = zero_vector(A.manifold, y)
    diff_group_apply!(A, tmp, Ie, y, F1)
    y2 = exp(A.manifold, y, tmp)
    F2 = dt * f(y2, sp, t + dt/2)
    diff_group_apply!(A, tmp, Ie, y, F2)
    exp!(base_manifold(A), y, y, tmp)
end
\end{lstlisting}

The presented method was applied to a case with two static charges: \SI{1e-6}{C} at position (\SI{-1}{m}, \SI{0}{m}) and \SI{-1e-6}{C} at position (\SI{1}{m}, \SI{0}{m}).
A rod of length \SI{0.67}{m} with charge density \SI{1e-6}{C/m} and mass \SI{1}{kg} started from three different initial positions.
The trajectories obtained in a simulation with time step \SI{0.001}{s} are presented in Figure~\ref{fig:Heun}.
Simulation times between \SI{30}{s} and \SI{135}{s} were chosen to have the rod remain relatively close to point charges.
Position and orientation of the rod is indicated at 20 equally spaced times, along with a densely sampled trajectory of the center of the rod.

>>>>

Code~\ref{lst:rod-dynamics} shows functions that calculate the right hand side of the ODE, where \lstinline{E_field} is a function that calculates electric field strength from the point charges at the given point.
The function \lstinline{force_and_torque} calculates the force and torque affecting the rod, while \lstinline{dynamics} converts them to the Lie algebra element can be used by the solver.
\begin{lstlisting}[
    language = Julia,
    numbers=left,
    label={lst:rod-dynamics},
    caption={Dynamics of a rod affected by point charges.},
    mathescape
]
function force_and_torque(
        sp::SystemParameters, Fθ, r
    )
    u, u${}_\perp$ = eachrow(Fθ)
    F, $\tau$ = zeros(2), 0.0
    for s in range(-sp.L/2, sp.L/2; length=Nq)
        E = E_field(sp, r .+ s .* u)
        F .+= sp.$\lambda$ * E * (sp.L/(Nq-1))
        $\tau$ += sp.$\lambda$ * s * dot(u${}_\perp$, E) * (sp.L/(Nq-1))
    end
    return F, $\tau$
end
# f((p,θ), (dp,dθ), t) including system parameters
function dynamics(state, sp::SystemParameters, t)
    p, dp = state.x
    F, tau = force_and_torque(sp, p.x...)
    a = F ./ sp.M
    coords = [tau / sp.Icm, a...]
    ddx = hat(sp.se2, coords, ArrayPartition)
    return ArrayPartition(dp, ddx)
end
\end{lstlisting}

@@

\begin{lstlisting}[
    language = Julia,
    numbers=left,
    label={lst:heun},
    caption={Step calculation for the RKMK Heun method.}
]
function integrator_step!(
        A::GroupAction, sp::SystemParameters,
        y, f, t, dt
) # The dynamics function is passed as f
    Ie = Identity(A.group)
    F1 = (dt / 2) * f(y, sp, t)
    tmp = zero_vector(A.manifold, y)
    diff_group_apply!(A, tmp, Ie, y, F1)
    y2 = exp(A.manifold, y, tmp)
    F2 = dt * f(y2, sp, t + dt/2)
    diff_group_apply!(A, tmp, Ie, y, F2)
    exp!(base_manifold(A), y, y, tmp)
end
\end{lstlisting}

The presented method was applied to a case with two static charges: \SI{1e-6}{C} at position (\SI{-1}{m}, \SI{0}{m}) and \SI{-1e-6}{C} at position (\SI{1}{m}, \SI{0}{m}).
A rod of length \SI{0.67}{m} with charge density \SI{1e-6}{C/m} and mass \SI{1}{kg} started from three different initial states:
\begin{enumerate}
    \item position (\SI{0}{m}, \SI{-1.5}{m}), angle \SI{0.1}{rad}, velocity (\SI{0}{m/s}, \SI{0}{m/s}), angular velocity \SI{0}{rad/s},
    \item position (\SI{-0.2}{m}, \SI{1}{m}), angle \SI{0.2}{rad}, velocity (\SI{-0.12}{m/s}, \SI{-0.15}{m/s}), angular velocity \SI{0}{rad/s},
    \item position (\SI{-0.5}{m}, \SI{-0.5}{m}), angle \SI{0.6}{rad}, velocity (\SI{-0.04}{m/s}, \SI{0.12}{m/s}), angular velocity \SI{0}{rad/s}.
\end{enumerate}
The trajectories obtained in a simulation with time step \SI{0.001}{s} are presented in Figure~\ref{fig:Heun}.
Simulation times between \SI{30}{s} and \SI{135}{s} were chosen to have the rod remain relatively close to point charges.
Position and orientation of the rod is indicated at 20 equally spaced times, along with a densely sampled trajectory of the center of the rod.
